---
title: "HW4 -Stefan Zdraljevic"
output: html_document
---


```{r}
library(scales)
library(dplyr)
```

#### load yeast cell cycle data

```{r}
yeast <- read.table("~/Dropbox/AndersenLab/LabFolders/Stefan/Courses/Stats_for_Bioinformatics/Homework/HW4/yst_cell_cycle.txt",
                    fill=T,header=T, sep="\t")
colnames(yeast) <- c("gene",colnames(yeast[,2:83]))
```

#### subset data to only contain cdc28 arrested experiments and remove NA genes

remove cdc28_140 timepoint because it eliminates a lot of data

```{r, echo=FALSE}
cdc28 <- yeast%>%
  select( contains("cdc28"))%>%
  select(-cdc28, -cdc28_140)
row.names(cdc28) <- yeast[,1]
cdc28 <- na.omit(cdc28)
```

#### cluster based on correlation distance

Little is gained from the gene clustering, but you can see that that certain timepoints cluster together due to proximity

```{r}
dd <- as.dist((1 - cor(t(cdc28),use = "pairwise.complete.obs"))/2)
hc=hclust(dd)
plot(hc, main = "HCL or Genes")

dd1 <- as.dist((1 - cor(cdc28,use = "pairwise.complete.obs"))/2)
hc1=hclust(dd1)
plot(hc1,main = "HCL or Timepoints")
```

#### principal component analysis and k-means clustering

```{r}
##############
##principle component analysis
eigenResult <- eigen(var(cdc28))     ## Eigen analysis
eigenVector <- eigenResult$vectors   ##extract Eigen vectors
p1 <- eigenVector[,1]                ## first eigen vector
p2 <- eigenVector[,2]                ## second eigen vector

cdc28m <- as.matrix(cdc28)
PCA1 <- cdc28m%*%p1                  ## convert to principal component 1
PCA2 <- cdc28m%*%p2                 ## convert to principal component 2
plot(PCA1,PCA2)
plot(eigenResult$values/sum(eigenResult$values))

View(eigenResult$values/sum(eigenResult$values))

varcomp <- data.frame(varexp = eigenResult$values/sum(eigenResult$values))
varcomp$expl <- cumsum(varcomp$varexp)
print(varcomp)
```

**NOTE ABOVE** 12 principal components are required to capture 95% variance in the data.

```{r}
KmsYeast <- kmeans(cdc28,12)
##
Yeast2 <- cbind(cdc28,PCA1,PCA2,KmsYeast$cluster)
plot(PCA1,PCA2,type="n", ylim = c(-3,3), xlim = c(-4,4))

for (i in 1:12){
  points(PCA1[Yeast2[,19]==i],PCA2[Yeast2[,19]==i],col=alpha(i,.5),pch=20)
}
```

## observe startification of k-means clusters on principal components space.









