---
title: 'Homework #2 - Stefan Zdraljevic'
output: html_document
---

# Normalization of Affymetrix data set 

This dataset was obtained from the GEO database - Reference series **GSE21784**.

The data consists of *C.elegans* expression data at three timepoints during the lifecycle (L4, adult day 6, adult day 15). Each timepoint is replicated in triplicate.

Data was downloaded, saved, and unzipped in my homeworke dirtectory

### Load required packages

```{r, warning=FALSE,message=FALSE}
# source("http://bioconductor.org/biocLite.R")
# biocLite("affy")
library(affy)

setwd("~/Northwestern/Courses/Stats_for_Bioinformatics/Homework/HW2/GSE21784_RAW/")
```

### Read in .CEL files and rename

```{r}
expd <- ReadAffy()
sampleNames(expd) <- c("L4_1","L4_2","L4_3", "Day6_1", "Day6_2","Day6_3","Day15_1","Day15_2","Day15_3")
```

### Plot two replicate experiments against each other

```{r,warning=FALSE}
y <- log2(pm(expd)[,1])[1:5000]
x <- log2(pm(expd)[,2])[1:5000]
LO <- loess(y~x,span=0.3)

plot(x,y,main="L4 vs L4 replicates",xlab="L4 replicate",ylab="L4 replicate")
points(x,predict(LO),col=2,lwd=.5,cex=.3)
```

### Plot L4 vs Day 6

```{r}
x <- log2(pm(expd)[,1])[1:5000]
y <- log2(pm(expd)[,4])[1:5000]
LO <- loess(y~x,span=0.3)

plot(x,y,main="L4 vs Day 6",xlab="Day 6 replicate",ylab="L4 replicate")
points(x,predict(LO),col=2,lwd=.5,cex=.3)
```

### Plot L4 vs Day 15

```{r}
x <- log2(pm(expd)[,1])[1:5000]
y <- log2(pm(expd)[,7])[1:5000]
LO <- loess(y~x,span=0.3)

plot(x,y,main="L4 vs Day 15",xlab="Day 15 replicate",ylab="L4 replicate")
points(x,predict(LO),col=2,lwd=.5,cex=.3)
```

### Histogram of log intensity

```{r}
hist(expd[,1:9],col=c(rep(2,3),rep(3,3),rep(4,3)), ylab = "Density", xlab = "Log Intensity")
```

### Boxplots of replicates

```{r}
boxplot(expd,col=c(rep(2,3),rep(3,3),rep(4,3)), xlab = "Stage", ylab = "Log Intensity")
```

### Identification of possible RNA degradation at different ends of the mRNA transcript

```{r}
deg <- AffyRNAdeg(expd[,1])
plotAffyRNAdeg(deg)  
```

### MVA plot of different day measurements

```{r}
gn <- sample(geneNames(expd),1000)       
pms <- pm(expd[,c(1,4,7)],gn)
mva.pairs(pms)
```

### Normalization of data

```{r, message=FALSE,warning=FALSE}
normalized.expd <- merge(normalize(expd[,1:3],method="loess"), 
                         normalize(expd[,4:6],method="loess"))

normalized.expd <- merge(normalized.expd, 
                         normalize(expd[,7:9],method="loess"))

boxplot(normalized.expd,col=c(rep(2,3),rep(3,3),rep(4,3)), main = "Normalized Expression Data", xlab = "Stage", ylab = "Log Expression")
```

### Normalized MVA compared to un-normalized

```{r}
gn <- sample(geneNames(normalized.expd),1000)   

pms <- pm(normalized.expd[,c(1,2,3)],gn)
mva.pairs(pms, main = "Normalzed MVA Plot")
     
pms <- pm(expd[,c(1,2,3)],gn)
mva.pairs(pms, main = "unNormalized MVA Plot")
```




